\chapter{Problem Definition}
\label{chapter01}

Tactical turn-based games like Duelyst \citep{duelyst} and Faeria \citep{faeria} have become very popular in the recent years. It is easy to create more content for such games (new spells, monsters, etc.) by simply changing the attributes of existing content. However, it is difficult to make sure the game stays balanced as the new content is added over time. This is generally done by human play-testing in-house before the game updates are released, and takes large amounts of time.

This thesis aims to solve the problem with \emph{procedural content generation}
(PCG)\@. We define PCG as `the algorithmic creation of game content with
little or no user input' \citep{pcgbook} We introduce a custom game similar to
Duelyst, and show that search based PCG can create new content that is
balanced.

\section{Types of PCG currently being used}

Many games utilize different kinds of PCG these days. Large commercial games
such as Diablo 2 \citep{diablo} use PCG to generate unique map layout of certain areas.
Some games like the popular Minecraft \citep{minecraft} go to an extreme and procedurally
generate the whole world. But PCG is not only limited to large games, even
small indie games like Spelunky \citep{spelunky} or Terraria \citep{terraria} utilize
PCG for generating playable levels.

Games also use PCG for item creation, where the specific attributes of an item
are generated online (e.g. Borderlands \citep{borderlands}).

\section{Thesis Goals}

Our main goal for this thesis is to generate balanced encounters. For this, we
decided to implement a custom game with flexible game mechanics so that there
are many different ways to create a balanced encounter. The game is turn-based,
zero-sum, with perfect information (see \autoref{chapter02}).

We implement the game both in the form of a simulator that can be used as a
library, and a GUI that a human player can use to play the game and test it. We
also implement an AI for the game so that we can automatically evaluate and
test games in our PCG algorithm.

In summary, the goals are:

\begin{itemize}
\item Design and implement a custom game, both GUI and a simulator of the game mechanics.
\item Build an AI that can be used to represent the player in terms of skill.
\item Create an algorithm for generating balanced encounters for our game.
\item Verify that the AI and encounter balancing works by conducting an experiment.
\end{itemize}

\section{Scope}

We narrow our scope to turn-based games, since simultaneous move games provide
an additional challenge when building an AI, which for us is just a means to
help with PCG\@. Our game zero-sum for two players. We also chose to make the
game with perfect information to simplify the creation of AIs.

\section{Used technology}

We chose \Csh{} and the .NET platform since it provides a good balance of programmer productivity and performance. \Csh{}, while being a high level language, still has the ability to control memory layout of objects to some extent. Most importantly, it has the \verb|struct| keyword which allows for objects which are allocated in-place and have copy semantics, contrast to Java, where all objects are dynamically allocated on the heap. This feature allows for compact representation of data and thus improves locality.

On the other hand, \Csh{} also provides high level primitives in the form of LINQ and the Task Parallel Library \citep{tpl} (see \autoref{threading-tpl}). An example of where this is used is projectiles, where one can create a projectile and \verb|await| the task that represents the projectile hitting the target, and chain an animation afterwards. While developing, we also made heavy use of the advanced profiling tools provided for the .NET ecosystem, most notably dotTrace \citep{dot-trace}.

